### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@985d7c950248429a9991142d4c0691e94823181b 🚀

RELOCATE: Indeed, it only assumes the action of well described
evolutionary mechanisms. The benefit of this approach is to provide the
potential for a more dynamic and nuanced understanding of risk: not only
on ecological timescales, but also by providing clues about which areas
can change over micro-evolutionary timescales.

Our study focuses largely on the biogeography of hosts. Specifically, we
identify the Amazon and South-Eastern Asia as hotspots where the
phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
(Anthony et al. 2017); surprisingly, current data suggest that viral
sharing between hosts is high in the Amazon and low in South-Eastern
Asia, which has the potential to result in different evolutionary
dynamics between these two regions, hinting at different futures for
their viral communities. This work is important both as a description of
the bats-betacoronaviruses complex, but also because more broadly, bats
are known reservoirs for a variety of emerging viruses and other
pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
the needs for bat conservation and disease prevention most likely very
difficult and a source of human-wildlife conflicts, especially in more
densely populated areas (Stone et al. 2015, Rego et al. 2015).

Yet, we know that viruses with high host plasticity, that is, the
ability of a given virus to adapt to various taxonomic orders and
ecological groups (Kreuder Johnson et al. 2015), are more likely to
amplify viral spillover, followed by secondary human-to-human
transmission, and geographical spread (Hazarie et al. 2021). High viral
host plasticity is an especially important trait for RNA viruses like
betacoronaviruses (Haddad et al. 2021). Indeed, our analysis of viral
sequences reveals that Latin America is a hotspot of viral
distinctiveness, suggesting that this part of the bats-betacoronaviruses
complex may be undergoing independent evolutionary dynamics (related
species sharing viruses that are different from the rest of the global
pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
richness is high but sharing is low; this suggests a different type of
evolutionary dynamics (unrelated viruses coevolving with evolutionarily
distinct hosts, generating high diversity locally, see \emph{e.g.}
Latinne et al. 2020). Both of these areas should be priority areas for
sampling, especially since Becker et al. (2022) advance that they harbor
undiscovered hosts of beta-coronaviruses. This diversity of hosts, and
the mechanisms by which the exchange of viruses occurs between species,
is largely affected by the local environmental conditions and
environmental change.

There are several factors that drive changes in the diversity of bats
(Alves et al. 2018), but human activities' effects on the ecosystem
(like modifications of land use) could significantly decrease it.
Therefore, it can be suggested that changes in the diversity of betacovs
in bats are linked to their biogeographic variation, and human
population density and other anthropogenic factors are decisive
moderators for its implications in public health. With the increase of
contact between humans and potential hosts, we also increase the risk of
emergence of novel diseases (Johnson et al. 2020), as previous studies
on RNA viruses suggest the importance of host phylogeography at the time
of virus dispersal (Gryseels et al. 2017). One of these scenarios where
interaction between bats and humans can occur can be seed dispersal in
tropical agroecosystems. It opens the discussion of whether the fruits
thrown by bats not only disperse seeds but could also be a source of
indirect interaction between viruses of bat origin and humans (Deshpande
et al. 2022). This represents a challenge for conservation strategies
and disease ecology since some areas can haveboth potential for the
acquisition of zoonotic viruses and bat-human interactions; in
particular, the challenge lies in the fact that actual exposure must
then be quantified accounting for several transmission scenarios,
including both direct and indirect bat - human interaction.

@@

\leavevmode\hypertarget{ref-Haddad2021SarPos}{}%
Haddad, D. et al. 2021. SARS-CoV-2: Possible recombination and emergence
of potentially more virulent strains (H Attoui, Ed.). - PLOS ONE 16:
e0251368.

@@

\leavevmode\hypertarget{ref-KreuderJohnson2015SpiPan}{}%
Kreuder Johnson, C. et al. 2015. Spillover and pandemic properties of
zoonotic viruses with high host plasticity. - Scientific Reports 5:
14830.

@@

\leavevmode\hypertarget{ref-Melaun2014BatPot}{}%
Melaun, C. et al. 2014. Bats as Potential Reservoir Hosts for
Vector-Borne Diseases. - In: Klimpel, S. and Mehlhorn, H. (eds), Bats
(Chiroptera) as Vectors of Diseases and Parasites: Facts and Myths.
Parasitology Research Monographs. Springer, pp. 25--61.

@@

\leavevmode\hypertarget{ref-Rego2015AssHum}{}%
Rego, K. M. da C. et al. 2015. Assessing human-bat interactions around a
protected area in northeastern Brazil. - Journal of Ethnobiology and
Ethnomedicine 11: 80.

@@

\leavevmode\hypertarget{ref-Stone2015ManCon}{}%
Stone, E. et al. 2015. Managing Conflict between Bats and Humans: The
Response of Soprano Pipistrelles (Pipistrellus pygmaeus) to Exclusion
from Roosts in Houses. - PLoS ONE 10: e0131825.

>>>>

Previous studies that have used host richness as proxy have predicted a
high diversity of unsampled bat viruses {[}Olival{]}, bat coronaviruses
{[}Anthony{]}, and even specifically betacoronaviruses {[}Becker{]} in
both the Amazon and southeast Asia. While we find that both regions are
characterized by highly divergent host and viral communities, our
framework identifies key differences between the regions. We find that
Latin America is a hotspot of both host and viral distinctiveness,
suggesting that this branch of the bat-betacoronavirus complex may be
undergoing independent evolutionary dynamics from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America {[}Anthony{]}. In contrast, in southeast
Asia, host richness and viral distinctiveness are high but sharing is
low; this suggests a different type of evolutionary dynamics that could
generate high viral diversity locally (see \emph{e.g.} (Latinne et al.
2020), as well as the discovery of recombinant viruses that share
genetic material from both the SARS-CoV and SARS-CoV-2 branches of the
Sarbecovirus lineage {[}Wu{]}). Both of these regions are priority areas
for sampling, especially given predictions by Becker et al. (2022) that
they contain many bat hosts of undiscovered betacoronaviruses. However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region {[}Carlson{]}.

{[}para about conservation{]} There are several factors that drive
changes in the diversity of bats (Alves et al. 2018), but human
activities' effects on the ecosystem (like modifications of land use)
could significantly decrease it. Therefore, it can be suggested that
changes in the diversity of betacoronaviruses in bats are linked to
their biogeographic variation, and human population density and other
anthropogenic factors are decisive moderators for its implications in
public health. With the increase of contact between humans and potential
hosts, we also increase the risk of emergence of novel diseases (Johnson
et al. 2020), as previous studies on RNA viruses suggest the importance
of host phylogeography at the time of virus dispersal (Gryseels et al.
2017). One of these scenarios where interaction between bats and humans
can occur can be seed dispersal in tropical agroecosystems. It opens the
discussion of whether the fruits thrown by bats not only disperse seeds
but could also be a source of indirect interaction between viruses of
bat origin and humans (Deshpande et al. 2022). This represents a
challenge for conservation strategies and disease ecology since some
areas can haveboth potential for the acquisition of zoonotic viruses and
bat-human interactions; in particular, the challenge lies in the fact
that actual exposure must then be quantified accounting for several
transmission scenarios, including both direct and indirect bat - human
interaction.

@@

@@

@@

@@

@@
